Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2014-10-14 Report/conclusions.tex
modified with markku input
Updated with rohit comments

In this paper, we addressed the allocation of space-frequency resources to the users in a multi-cell \ac{MIMO} \ac{OFDM} system. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user can be achieved by a non-zero precoding vector. We proposed the \ac{JSFRA} scheme by adopting the \ac{SCA} technique to model the nonconvex constraint as a sequence of convex subsets to design the precoders for minimizing the number of queued packets. Additionally, an alternative approach using \ac{MSE} relaxation is also proposed with the same objective by fixing the receivers based on \ac{MSE} minimization. We also proposed the distributed solutions for the centralized \ac{JSFRA} problem using \ac{ADMM} and primal decomposition methods. Finally, we proposed an iterative algorithm to determine the precoders in a decentralized manner based on the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme. Numerical results shows that the proposed algorithms perform better than the existing approaches.

>>>>

In this paper, we addressed the allocation of space-frequency resources to the users in a multi-cell \acl{MIMO} \acl{OFDM} system. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user can be achieved by a non-zero precoding vector. We proposed the \ac{JSFRA} scheme by adopting the \ac{SCA} technique to model the nonconvex constraint as a sequence of convex subsets to design the precoders for minimizing the number of queued packets. Additionally, an alternative approach using \acl{MSE} relaxation is also proposed with the same objective by fixing the receivers based on \ac{MSE} minimization. We also proposed the distributed solutions for the centralized \ac{JSFRA} problem using primal and \ac{ADMM} decomposition methods. Finally, we proposed an iterative algorithm to determine the precoders in a decentralized manner based on the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme. Numerical results shows that the proposed algorithms perform better than the existing approaches.

### 2014-10-14 Report/simulation_distributed.tex
Updated with rohit comments

Fig. \ref{fig-d-2} plots the performance of the centralized and the distributed algorithms at each \ac{SCA} update. In case of the distributed approaches, in between each \ac{SCA} update, the primal or the \ac{ADMM} exchanges are performed for \me{J_{\max} = 20} iterations. In practice, \me{J_{\max} = 1} can be set to perform the \ac{SCA} update, \ac{ADMM} or primal update, and the receive beamformers \me{\mvec{W}{k,n}} update at the same instant, which affects the convergence rate and the stationary point of the final solution. The data tips are used to highlight the convergent points of various algorithms. The performance of the \ac{JSFRA} schemes using the primal decomposition are notably inferior compared to the \ac{ADMM} approach for the same schemes. It is mainly attributed to the difficulty in selecting the step size for the system employing \me{N_B \geq 3} \acp{BS}.
\begin{figure*}
	\centering
	\subfloat[][System \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 3,2,8,1 \rbrace}]{
		\includegraphics[width=0.48\textwidth]{fig-3-2}
		\label{fig-d-1}
	}
	\hfill
	\subfloat[][System \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 6,3,12,2 \rbrace}]{
		\includegraphics[width=0.48\textwidth]{fig-5-1}
		\label{fig-d-2}
	}
	\caption{Number of backlogged packets at each \ac{SCA} points}
	\label{fig-d}
\end{figure*}
% \begin{figure*}
% \centering
% \begin{subfigure}{0.49\textwidth}
% \includegraphics[width=\textwidth]{fig-5}
% \caption{Queue deviation}
% \end{subfigure}
% \hfill
% \begin{subfigure}{0.49\textwidth}
% \includegraphics[width=\textwidth]{fig-6}
% \caption{Sum rate performance}
% \end{subfigure}
% \caption{Convergence plot for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 6,3,12,2 \rbrace} model}
% \label{fig-d-2}
% \end{figure*}
}{
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{fig-3-2}
\caption{Convergence behaviour of the centralized and the distributed algorithms for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 3,2,8,1 \rbrace}}
\label{fig-d} \vspace{-0.1in}
\end{figure}
}
Fig. \ref{fig-d-3.1} compares the performance of the centralized and the \ac{KKT} algorithm in Section \ref{sec-4.3} for different exponents. The plots present the surplus number of backlogged packets at the end of each iteration or the \ac{SCA} update point. Fig. \ref{fig-d-3.1} shows that the \me{\ell_1} norm for \ac{JSFRA} scheme provides better performance over the rest of the schemes due to its greedy objective. The \ac{KKT} approach for \me{\ell_1} norm is not defined due to the non-differentiability of the objective as discussed in the Section \ref{sec-4.3}, thereby performs the worst of all other approaches. The heuristic method used in the figure is obtained by forcing the dual variable \me{\sigma_{l,k,n}} in \eqref{kkt-mse-4.2} to \me{0} when the queue deviation is negative \me{Q_k - t_k < 0}, if not, then it will be the same as in \eqref{kkt-mse-4.2}. The additional condition can be justified due to the dropping of absolute value operator from the objective. It can be seen that the heuristic method oscillates near the optimal point with the deviation determined by the factor \me{\rho} used in \eqref{kkt-mse-4.1}.
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{fig-9-3}
	\caption{Impact of varying \me{q} in the total number of backlogged packets after each \ac{SCA} update for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
	\label{fig-d-3.1} \vspace{-0.1in}
\end{figure}

The objective values are mentioned in the legend for all the schemes, since the objective of the \me{\ell_2} norm is not the same as that of the \me{\ell_1} norm, which is used for the plot. The \me{\ell_2} norm for the \ac{JSFRA} and the \ac{KKT} based approach achieves nearly the same objective value of \me{6.62} but different \me{\chi}, since the dual variables \me{\alpha_{l,k,n}} and \me{\sigma_{l,k,n}} are not iterated until convergence between each \ac{SCA} steps in the \ac{KKT} approach. In the simulation, we update all the variables at once at each iteration, \textit{i.e}, \me{J_{\max} = 1}, which is well justified for the practical implementations due to the signaling overheads. Fig. \ref{fig-d-3.1} also compares the effect of dropping the squared rate variable from the objective in the \ac{Q-WSRME} scheme compared to the \me{\ell_2} norm which includes it. By dropping the squared rate variable, the \ac{Q-WSRME} scheme minimizes the number of queued packets in a prioritized manner based on the respective queues. On contrary, the \me{\ell_2} norm allocate rates to the users with the higher number of queued packets before addressing the users with the smaller number of queued packets.

>>>>

Fig. \ref{fig-d-2} plots the performance of the centralized and the distributed algorithms at each \ac{SCA} update. In case of the distributed approaches, in between each \ac{SCA} update, the primal or the \ac{ADMM} exchanges are performed for \me{J_{\max} = 20} iterations. In practice, \me{J_{\max} = 1} can be set to perform the \ac{SCA} update, \ac{ADMM} or primal update, and the receive beamformers \me{\mvec{W}{k,n}} update at the same instant, which affects the convergence rate and the stationary point of the final solution. The data tips are used to highlight the convergent points of various algorithms. The performance of the \ac{JSFRA} schemes using the primal decomposition are notably inferior compared to the \ac{ADMM} approach for the same schemes. It is mainly attributed to the difficulty in selecting the step size for the system employing \me{N_B \geq 3} \acp{BS}.
\begin{figure*}
	\centering
	\subfloat[][System \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 3,2,8,1 \rbrace}]{
		\includegraphics[width=0.48\textwidth]{fig-3-2}
		\label{fig-d-1}
	}
	\hfill
	\subfloat[][System \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 6,3,12,2 \rbrace}]{
		\includegraphics[width=0.48\textwidth]{fig-5-1}
		\label{fig-d-2}
	}
	\caption{Number of backlogged packets at each \ac{SCA} points}
	\label{fig-d}
\end{figure*}
% \begin{figure*}
% \centering
% \begin{subfigure}{0.49\textwidth}
% \includegraphics[width=\textwidth]{fig-5}
% \caption{Queue deviation}
% \end{subfigure}
% \hfill
% \begin{subfigure}{0.49\textwidth}
% \includegraphics[width=\textwidth]{fig-6}
% \caption{Sum rate performance}
% \end{subfigure}
% \caption{Convergence plot for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 6,3,12,2 \rbrace} model}
% \label{fig-d-2}
% \end{figure*}
}{
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{fig-3-2}
\caption{Convergence behaviour of the centralized and the distributed algorithms for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 3,2,8,1 \rbrace}}
\label{fig-d} \vspace{-0.1in}
\end{figure}
}
Fig. \ref{fig-d-3.1} compares the performance of the centralized and the \ac{KKT} algorithm in Section \ref{sec-4.3} for different exponents by plotting the total number of backlogged packets at each \ac{SCA} update point. The \me{\ell_1} norm \ac{JSFRA} scheme provides better performance over other schemes due to the greedy objective. The \ac{KKT} approach for \me{\ell_1} norm is not defined due to the non-differentiability of the objective as discussed in the Section \ref{sec-4.3}. If used for \me{\ell_1} norm, the problem of over-allocation will not affect the dual variables \me{\sigma_{l,k,n}} and \me{\alpha_{l,k,n}} since the queue deviation is raised to the power zero in \eqref{kkt-mse-4.2}, which will always be equal to one. A heuristic method is proposed in Fig. \ref{fig-d-3.1} by assigning zero to the dual variable \me{\sigma_{l,k,n}} when the queue deviation is negative, \textit{i.e}, \me{Q_k - t_k < 0}. This additional condition is required to address the problem of over-allocation in the \me{\ell_1} norm for dropping the absolute value operator from the objective function. It can be seen that the heuristic method oscillates near the optimal point with the deviation determined by the factor \me{\rho} used in \eqref{kkt-mse-4.1}.
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{fig-9-3}
	\caption{Impact of varying \me{q} in the total number of backlogged packets after each \ac{SCA} update for a system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
	\label{fig-d-3.1} \vspace{-0.1in}
\end{figure}

The objective values are mentioned in the legend for all the schemes and the objective of the \me{\ell_2} norm is not the same as that of the \me{\ell_1} norm used for plotting. For simulations, we update all variables in \eqref{kkt-mse-4} at once at each iteration, \textit{i.e}, \me{J_{\max} = 1}, which is well justified for the practical implementations due to the signaling overheads. The \me{\ell_2} norm for the \ac{JSFRA} and the \ac{KKT} approach achieves nearly the same value of \me{6.62} with different \me{\chi}, due to the limited number of iterations for the dual variable convergence between each \ac{SCA} update. Fig. \ref{fig-d-3.1} also shows the effect of dropping the squared rate variable from the objective in the \ac{Q-WSRME} scheme compared to the \me{\ell_2} norm which includes it. By dropping it, the \ac{Q-WSRME} scheme minimizes the number of queued packets in a prioritized manner based on the respective queues. On contrary, the \me{\ell_2} norm allocate rates to the users with the higher number of queued packets before addressing the users with the smaller number of queued packets.
